Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2015-11-10 Publications/2015-NeuroImage/responseLetter2.tex

\bigskip
\bigskip
\subsection*{Reviewer \#1:}
\begin{reviewpoint}
The paper describes an approach for the delineation of selected anatomical structures ([...]) in multi-shell diffusion MR images
  by registering a segmented T1 weighted reference image with fractional anisotropy and apparent diffusion coefficient maps derived
  from the diffusion weighted images.

The approach corresponds fully to the usual registration of an atlas to the image to be segmented, leaving me rather clueless why do
  they call the method ``simultaneous registration and segmentation''.

\end{reviewpoint}
\reply{%
The description of the method given by both reviewers (see \autoref{com:10}, Reviewer \#2) lead us to think that the paper
  did not properly describe the actual inputs and outputs the proposed methodology.
In general, we have almost-completely reformulated the abstract, introduction and discussion to define this
  point clearly.
Particularly, the following sections have been introduced to give an answer to this particular comment:
\begin{itemize}
\item Lines 45--53 (p. 2).
\item Lines 244--246 (p. 13), here we introduce the previously missing discussion to call the
  method ``simultaneous registration and segmentation''.
\item Lines 250--252 (p. 13), we comment here on the similarity of \emph{regseg} to the atlas-based segmentation
  using joint registration and segmentation methods.
\end{itemize}
}

\begin{reviewpoint}
The paper describes an approach [...] (quite some are listed as possible examples but results are only reported for the white matter,
  deep gray matter and the pial surface) in multi-shell [...].
\end{reviewpoint}
\reply{
To clarify this point, we have stated more clearly that the choice of surfaces to segment the target data is
  crucial (lines 46--50, p.2).
We have better defined the segmentation model that drives the registration process (lines 94--95, p. 5).
We have also improved the former Figure 4 (now it is Figure 2, page 5), and brought it earlier in the paper to
  illustrate earlier the segmentation model.
Since the former visualization of the conditional (region-wise) joint distribution of FA and ADC was a bit clumsy,
  we have extracted the conditional distributions to the right side of each panel.
We expect that these modifications make clear the difference between the surfaces involved in driving the registration
  process and the surfaces in which we evaluate the error of the method (lines 217--218, p. 10) are a subset of the
  former.
}

\begin{reviewpoint}
The proposed algorithm is rather standard, relying on a region-based elastic contour deformation approach.
The cost function is derived in the usual Bayesian framework, relying on a multivariate normal distribution prior for the image
  intensities and a Tikhonov regularizer enforcing a predefined normal distribution over the distortions along the imaging axes.
In this respect the novelty of the paper can be questioned, while the application context and the very promising results may still
  justify a publication in NeuroImage.
\end{reviewpoint}
\reply{
We agree with the reviewer on that the novelty of the methodological elements of our approach is limited.
However, to our knowledge there exists no alternative method combining the same elements to segment
  multivariate data.
We have moved the related methods and our methodological contributions from the Introduction to a more appropriate
  place in the Discussion (lines 261--278), additionally satisfying \autoref{com:9} (Reviewer \#2) suggesting to include
  in the introduction only the references related to the focus of the paper.
The particular changes can be listed as follows:
\begin{itemize}
  \item Lines 252--260 (p. 13), we describe the methodological differences w.r.t. \cite{gorthi_active_2011},
    one of the latest atlas-based segmentation methods using simultaneous registration and segmentation.
  \item Lines 261--270 (p. 13), we compare the methodological features of our method and \emph{bbregister},
    an antecedent and standardized tool used on the same application.
  \item Lines 271--273 (p. 13), we introduce the methodological differences with a recent alternative method in
    terms of regularization.
  \item Lines 273--278 (p. 13) mention some other minor methodological contributions.
\end{itemize}
}

@@

\begin{reviewpoint}
5. The vertex-wise error distributions are not only skewed, but quite clearly structured im many cases
  (especially for the phantom images).
What is the underlying reason?
\end{reviewpoint}
\reply{
We have added a discussion on this in the corresponding section, lines 287-293 (p. 14).
Thanks for pointing this systematic error out.
}

\bigskip
\bigskip
\subsection*{Reviewer \#2:}
\begin{reviewpoint}
This paper seems to present a method that can perform wM-GM segmentation and simultaneously estimate a deformation field to
  correct for EPI distortion.
Apparently, the idea is to use nested WM-GM surfaces, as well as other surfaces, to provide better cortical registration than
  a competing method, which is not clearly described.
The overall goal and motivation of the paper are not clearly explained, and almost all related references for EPI distortion
  correction are missing.
It is unclear if the deformation field solution is restricted to the phase encode direction.
It is possible that this is a very useful method, as the results seem to indicate, but the paper is unfortunately too unclear to
  make a full judgement about the method.
\end{reviewpoint}
\reply{
We thank the reviewer for pointing out that the overall goal is not clearly explained.
We have rewritten the abstract, introduction and discussion sections almost fully.
Particular comments included within this general comment are addressed later: \autoref{com:18}
  as regards EPI distortions, and \autoref{com:20} for the phase-encoding direction point.
In order to better focus the application field, we have introduced lines 11--14 (p. 2) and lines
  312--317 (p. 14).
We have also added potential applications in the new lines 331--340 (p. 15) to better illustrate
  the method.
}

\begin{reviewpoint}
abstract: Not immediately clear what this means: ``active contours without edges extracted from structural images''
  (Where do the edges come from? What modalities are actually input to regseg?)
\end{reviewpoint}
\reply{%
In our adaptation of ``active contours without edges'' \citep{chan_active_2001}, the active contours are
  ``extracted from structural images''.
The particular sentence has been removed from the abstract, and we expect that the comprehensive reformulation
  of the descriptive sections of the paper give a successful answer to this issue.
Regarding the actual inputs of \emph{regseg}, we have introduced modifications in the abstract, and in lines
  45--50 (p. 2) and lines 243--244 (p. 13).
}

@@

}

\begin{reviewpoint}
Please define what you mean here by segmentation.
It seems there is an assumption that it means specifically segmentation via registration, where the
  structures are defined using T1 images, but that is not stated anywhere.

Now on line 19 it seems the goal of the segmentation may simply be the segmentation of the WM.
So far, I have no idea what the overall goal of the approach is.
\end{reviewpoint}
\reply{
This comment is very related to \autoref{com:14}.
We have modified the abstract and introduction accordingly, especially lines 15--27 (p. 2).
}

\begin{reviewpoint}
Line 24. clustering of what?
\end{reviewpoint}
\reply{
This comment is a consequence of the two previous comments.
The reviewer is right in the sense that we did not clearly state what we are segmenting, and therefore
  the references in the former line 24 were a bit orphaned.
All methods reviewed in lines 15--27 cluster/segment/classify \emph{voxels} in dMRI space based on
  one or another feature derived from dMRI data (lines 18--20, p. 2).
We have clarified the manuscript according to \autoref{com:14} and \autoref{com:15} and this comment.
}

@@

\begin{reviewpoint}
Was the warp restricted to the phase encode direction?
If so, the r in eq (1) should be restricted to this direction.
I do not see this mentioned anywhere near the equations in the Methods section.
If the solution found for the deformation field was not restricted to the phase encode direction,
  then this is not a reasonable way to perform EPI distortion correction. Please clarify.

[...]

I still see nothing in the results indicating that the estimated distortions were restricted to the phase encode direction.
If not, the added degrees of freedom might result in improved surface segmentation, but the distortion field would not correspond at
  all to the physical problem under study.
Furthermore, the dMRI gradients would have to be re-oriented independently at each voxel, and this is not currently possible in any
  software I know of.
\end{reviewpoint}
\reply{
The general functionality of \emph{regseg} does not restrict distortion if it is not requested using the
  feature of anisotropic regularization.
For the experiments on real data, the ground-truth deformations and the deformation field estimated by \emph{regseg}
  are restricted to be nonzero only in the phase-encoding direction.
To make this more clear, we have introduced the following edits:
\begin{itemize}
  \item Lines 131--134 (p. 7).
  \item Lines 198--199 (p. 10).
\end{itemize}
}

\begin{reviewpoint}
The experiments seem reasonable and clearly presented in general. But far too little detail is given on the B0-T2 comparison method.
\end{reviewpoint}
\reply{
We thank the reviewer for positively weighing our experiments and results.
As stated in line 211 (p. 10), we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
  internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
A footnote in line 212 (p. 10) directly links the settings for this method.
}

\begin{reviewpoint}
Please define the phrase ``active contours without edges''.
\end{reviewpoint}
\reply{
Additionally to the edits in \autoref{com:10}, we have more clearly defined the concept with the discussion
  introduced in lines 267--270 (p. 13).
}

### 2015-11-10 Publications/2015-NeuroImage/06-discussion.tex
suggestions from AS

\revcomment[C9 (R.2)]{%
Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
  where there are precise surfaces delineating the structure, a target multivariate
  image in which the surfaces must be fitted, and the mapping between the surfaces and
  the volume encodes relevant physiological information, such as the normal/abnormal
  development or the macroscopic dynamics of organs and tissues.
For instance, \regseg{} may be applied in fields like neonatal brain image segmentation
  in longitudinal MRI studies of the early developmental patterns \citep{shi_neonatal_2010}.
In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
  propagated to the initial timepoints, regardless the changes of the contrast and spatial
  development between them.
More generally, \regseg{} may also be applied to the personalized study of longitudinal alteration
  of the brain using multispectral images, for instance in the case of traumatic brain
  injury \citep{irimia_structural_2014} or in the monitorization of brain tumors
  \citep{weizman_semiautomatic_2014}.}

>>>>

\revcomment[C9 (R.2)]{%
Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
  where there are precise surfaces delineating the structure, a target multivariate
  image in which the surfaces must be fitted, and the mapping between the surfaces and
  the volume encodes relevant physiological information, such as the normal/abnormal
  development or the macroscopic dynamics of organs and tissues.
For instance, \regseg{} may be applied in fields like neonatal brain image segmentation
  in longitudinal MRI studies of the early developmental patterns \citep{shi_neonatal_2010}.
In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
  propagated to the initial timepoints, regardless of the changes in the contrast and spatial
  development between them.
More generally, \regseg{} may also be applied to the personalized study of longitudinal alteration
  of the brain using multispectral images, for instance in the case of traumatic brain
  injury \citep{irimia_structural_2014} or in the monitorization of brain tumors
  \citep{weizman_semiautomatic_2014}.}
